Adaptive Hybrid Screening for Efficient Lasso Optimization

### 2021-01-29 jscs/adaptive-hssr.tex
Moving journal files to jscs directory

@@

\begin{figure}[ht]
    \centering
    \includegraphics[width=0.6\textwidth]{plots/batchsizes.eps}    \caption{Proportion of features rejected by screening rules vs index of $\lambda$-values for different update delay.}
    \label{fig:batchsizes}
\end{figure}

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=\textwidth]{plots/511.eps}    \caption{Comparing speed of screening methods for lasso model under different settings. Top row: computation time in second. Bottom row: relative computation time compared to AC. First column: varying number of features $p$. Second column: varying sample size $n$. Third column: varying signal strength $\beta_{\max}$. Fourth column: varying sparsity $s$.}
    \label{fig:5.1.1a}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[width=0.82\textwidth]{plots/511b.eps}    \caption{Comparing speed of screening methods for lasso model under different settings. Top row: computation time in second. Bottom row: relative computation time compared to AC. First column: Varying number of grid points $L$. Second column: Varying range of grid $\lambda_{\min}/\lambda_{\max}$. Third column: Varying number of grid points $L$ with fixed spacing.}
    \label{fig:5.1.1b}
\end{figure}

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=\textwidth]{plots/512.eps}    \caption{Comparing speed of screening methods for sparse logistic regression model under different settings. Top row: computation time in second. Bottom row: relative computation time compared to SSR. First column: varying number of features $p$. Second column: varying sample size $n$. Third column: varying signal strength $\beta_{\max}$. Fourth column: varying sparsity $s$.}
    \label{fig:5.1.2a}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[width=0.82\textwidth]{plots/512b.eps}    \caption{Comparing speed of screening methods for sparse logistic regression model under different settings. Top row: computation time in second. Bottom row: relative computation time compared to SSR. First column: Varying number of grid points $L$. Second column: Varying range of grid $\lambda_{\min}/\lambda_{\max}$. Third column: Varying number of grid points $L$ with fixed spacing.}
    \label{fig:5.1.2b}
\end{figure}

@@

>>>>

\documentclass[]{interact}
\usepackage{epstopdf}% To incorporate .eps illustrations using PDFLaTeX, etc.
\usepackage[caption=false]{subfig}% Support for small, `sub' figures and tables
%\usepackage[nolists,tablesfirst]{endfloat}% To `separate' figures and tables from text if required
%\usepackage[doublespacing]{setspace}% To produce a `double spaced' document if required
%\setlength\parindent{24pt}% To increase paragraph indentation when line spacing is doubled

\usepackage[numbers,sort&compress]{natbib}% Citation support using natbib.sty
\bibpunct[, ]{[}{]}{,}{n}{,}{,}% Citation support using natbib.sty
\renewcommand\bibfont{\fontsize{10}{12}\selectfont}% Bibliography support using natbib.sty
\makeatletter% @ becomes a letter
\def\NAT@def@citea{\def\@citea{\NAT@separator}}% Suppress spaces between citations using natbib.sty
\makeatother% @ becomes a symbol again

\theoremstyle{plain}% Theorem-like structures provided by amsthm.sty
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

\theoremstyle{remark}
\newtheorem{remark}{Remark}
\newtheorem{notation}{Notation}

\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[ruled,vlined, linesnumbered]{algorithm2e}
\usepackage[english]{babel}
\usepackage[nottoc]{tocbibind}
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage{float}
\usepackage{hyperref}
\hypersetup{colorlinks=true, citecolor=black, linkcolor=., urlcolor=cyan}
\graphicspath{{../plots/}}

% Custom commands / shortcuts
\providecommand{\sign}{\textrm{sign}}
\providecommand{\pb}[1]{\textcolor{red}{#1}}
\providecommand{\lam}{\lambda}
\newcommand{\quotes}[1]{`#1'}

\title{Adaptive Hybrid Screening for Efficient Lasso Optimization}

\author{Word count: 8899.}

\date{}

\begin{document}

\maketitle

\begin{abstract}
Lasso type models are popular in statistics, especially for high-dimensional data. Typical practice is to tune the size of lasso penalty along a path of values. Due to modern data collection techniques, researchers need to analyze data with millions of features, which brings a great need for an efficient lasso optimization algorithm. Feature screening techniques have proven to be powerful for this need, because they can discard features and lead to a model with much less features. In this paper, we develop an adaptive hybrid screening framework where screening is carried out adpatively along the path of tuning parameter values. It reuses previous solutions in the path and reduces unhelpful heavy computations. It is a flexible framework that can be easily extended to different types of lasso models. Two applications for the standard lasso model and the sparse logistic model are shown as examples. We perform simulation study and real data study in a wide range of scenarios and the adaptive hybrid methods outperform other state-of-the-art method significantly and uniformly, with the greatest speedup in the most challenging scenarios.
\end{abstract}

\begin{keywords}
Lasso; feature Screening; safe screening; solution path; large-scale sparse learning
\end{keywords}

@@

\begin{figure}[ht]
    \centering
    \includegraphics[width=0.6\textwidth]{batchsizes.eps}    \caption{Proportion of features rejected by screening rules vs index of $\lambda$-values for different update delay.}
    \label{fig:batchsizes}
\end{figure}

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=\textwidth]{511.eps}    \caption{Comparing speed of screening methods for lasso model under different settings. Top row: computation time in second. Bottom row: relative computation time compared to AC. First column: varying number of features $p$. Second column: varying sample size $n$. Third column: varying signal strength $\beta_{\max}$. Fourth column: varying sparsity $s$.}
    \label{fig:5.1.1a}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[width=0.82\textwidth]{511b.eps}    \caption{Comparing speed of screening methods for lasso model under different settings. Top row: computation time in second. Bottom row: relative computation time compared to AC. First column: Varying number of grid points $L$. Second column: Varying range of grid $\lambda_{\min}/\lambda_{\max}$. Third column: Varying number of grid points $L$ with fixed spacing.}
    \label{fig:5.1.1b}
\end{figure}

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=\textwidth]{512.eps}    \caption{Comparing speed of screening methods for sparse logistic regression model under different settings. Top row: computation time in second. Bottom row: relative computation time compared to SSR. First column: varying number of features $p$. Second column: varying sample size $n$. Third column: varying signal strength $\beta_{\max}$. Fourth column: varying sparsity $s$.}
    \label{fig:5.1.2a}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[width=0.82\textwidth]{512b.eps}    \caption{Comparing speed of screening methods for sparse logistic regression model under different settings. Top row: computation time in second. Bottom row: relative computation time compared to SSR. First column: Varying number of grid points $L$. Second column: Varying range of grid $\lambda_{\min}/\lambda_{\max}$. Third column: Varying number of grid points $L$ with fixed spacing.}
    \label{fig:5.1.2b}
\end{figure}

@@

\renewcommand\bibname{References}
\bibliographystyle{tfnlm}
\begin{thebibliography}{10}
\providecommand{\url}[1]{\normalfont{#1}}
\providecommand{\urlprefix}{Available from: }

\bibitem{tibshirani1996regression}
Tibshirani~R. Regression shrinkage and selection via the lasso. Journal of the
  Royal Statistical Society: Series B (Methodological).
  1996;\hspace{0pt}58(1):267--288.

\bibitem{yuan2006model}
Yuan~M, Lin~Y. Model selection and estimation in regression with grouped
  variables. Journal of the Royal Statistical Society: Series B (Statistical
  Methodology). 2006;\hspace{0pt}68(1):49--67.

\bibitem{zou2005regularization}
Zou~H, Hastie~T. Regularization and variable selection via the elastic net.
  Journal of the Royal Statistical Society: Series B (Statistical Methodology).
  2005;\hspace{0pt}67(2):301--320.

\bibitem{friedman2007pathwise}
Friedman~J, Hastie~T, H{\"o}fling~H, et~al. Pathwise coordinate optimization.
  The Annals of Applied Statistics. 2007;\hspace{0pt}1(2):302--332.

\bibitem{Tibshirani2012}
Tibshirani~R, Bien~J, Friedman~J, et~al. Strong rules for discarding predictors
  in lasso-type problems. Journal of the Royal Statistical Society Series B.
  2012;\hspace{0pt}74(2):245--266.

\bibitem{qian2019fast}
Qian~J, Du~W, Tanigawa~Y, et~al. A fast and flexible algorithm for solving the
  lasso in large-scale and ultrahigh-dimensional problems. BioRxiv.
  2019;\hspace{0pt}:630079.

\bibitem{ghaoui2010safe}
Ghaoui~LE, Viallon~V, Rabbani~T. Safe feature elimination for the lasso and
  sparse supervised learning problems. arXiv:10094219. 2010;\hspace{0pt}.

\bibitem{wang2013lasso}
Wang~J, Zhou~J, Wonka~P, et~al. Lasso screening rules via dual polytope
  projection. In: Advances in Neural Information Processing Systems; 2013. p.
  1070--1078.

\bibitem{xiang2012fast}
Xiang~ZJ, Ramadge~PJ. Fast lasso screening tests based on correlations. In:
  International Conference on Acoustics, Speech and Signal Processing; IEEE;
  2012. p. 2137--2140.

\bibitem{xiang2011learning}
Xiang~ZJ, Xu~H, Ramadge~PJ. Learning sparse representations of high dimensional
  data on large scale dictionaries. In: Advances in Neural Information
  Processing Systems; 2011. p. 900--908.

\bibitem{Zeng2021}
Zeng~Y, Yang~T, Breheny~P. Hybrid safe–strong rules for efficient
  optimization in lasso-type problems. Computational Statistics \& Data
  Analysis. 2021;\hspace{0pt}153:107063.

\bibitem{zeng2017biglasso}
Zeng~Y, Breheny~P. The biglasso package: a memory-and computation-efficient
  solver for lasso model fitting with big data in {R}. arXiv:170105936.
  2017;\hspace{0pt}.

\bibitem{lee2007efficient}
Lee~H, Battle~A, Raina~R, et~al. Efficient sparse coding algorithms. In:
  Advances in neural information processing systems; 2007. p. 801--808.

\bibitem{Fan2008}
Fan~J, Lv~J. Sure independence screening for ultrahigh dimensional feature
  space. Journal of the Royal Statistical Society Series B.
  2008;\hspace{0pt}70(5):849--911.

\bibitem{wang2014safe}
Wang~J, Zhou~J, Liu~J, et~al. A safe screening rule for sparse logistic
  regression. In: Advances in Neural Information Processing Systems; 2014. p.
  1053--1061.

\bibitem{breheny2015group}
Breheny~P, Huang~J. Group descent algorithms for nonconvex penalized linear and
  logistic regression models with grouped predictors. Statistics and Computing.
  2015;\hspace{0pt}25(2):173--187.

\bibitem{xiang2016screening}
Xiang~ZJ, Wang~Y, Ramadge~PJ. Screening tests for lasso problems. IEEE
  Transactions on Pattern Analysis and Machine Intelligence.
  2016;\hspace{0pt}39(5):1008--1027.

\bibitem{Das2016}
Das~S, Forer~L, Sch{\"o}nherr~S, et~al. Next-generation genotype imputation
  service and methods. Nature Genetics. 2016 Oct;\hspace{0pt}48(10):1284--1287.

\end{thebibliography}

\end{document}
